---
title: "eda"
output: github_document
date: "2023-09-28"
---

first, load the libraries:

```{r}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```

second, load the same code as yesterday to pull the data: 

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) |>
  select(name, id, everything())
```

^ same code as 10/3, but added `month = lubridate` code 

## Initial numeric work

```{r}
weather_df |> 
  ggplot(aes(x = prcp)) + 
  geom_histogram()
```

here are the big outliers: 

```{r}
weather_df |> 
  filter(prcp > 1000)
```

```{r}
weather_df |> 
  filter(tmax >= 20, tmax <= 30) |> 
  ggplot(aes(x= tmin, y = tmax, color = name)) + 
  geom_point()
```

^ why does Molokai appear in bands? take a closer look at just the data first, important to understand what's going on with the data before experimenting

## Grouping

```{r}
weather_df |> 
  group_by(name, month)
```

^ 72 unique groups (of name, month) in this dataset 

## Counting

```{r}
weather_df |> 
  group_by(name, month) |> 
  summarize(n_obs = n())
```
^ the only things that have been retained are the variables defined by `group_by` structure, and counting the number of observations `n`

```{r}
weather_df |> 
  group_by(month) |> 
  summarize(n_obs = n())
```

```{r}
weather_df |> 
  count(name, name = "n_obs")
```
^ `count` works instead of `summarize` for a total # of observations instead of a list. name the column using `name =` 

```{r}
weather_df |> 
  count(name, month) |> 
  pivot_wider(
    names_from = name,
    values_from = n
  )
```

^ `pivot_wider` is the opposite purpose of `pivot_longer` -> make data table more human readable (`pivot_longer` makes data usable)


## General summaries

```{r}
weather_df |> 
  drop_na(tmax) |> 
  group_by(name, month) |> 
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE),
    median_tmax = median(tmax, na.rm = TRUE),
    sd_tmax = sd(tmax, na.rm = TRUE)
  )
```

^ table for what is the average tmax by name and month, by mean, median, and max for `tmax`. 

?? what is `na.rm`?? 

```{r}
weather_df |> 
  group_by(name, month) |> 
  summarize(mean_tmax = mean(tmax, na.rm = TRUE)) |> 
  ggplot(aes(x = month, y = mean_tmax, color=name)) +
  geom_point() +
  geom_line()
```

^gives a summary of the data - you can build off of a dataframe to  `mutate` or `group_by` once you have the dataframe. 

### pivot_wider and why it works

```{r}
weather_df |> 
  group_by(name, month) |> 
  summarize(mean_tmax = mean(tmax, na.rm = TRUE)) |> 
  pivot_wider(
    names_from= name,
    values_from = mean_tmax
  ) |> 
  knitr::kable(digits = 2)
```

^ make a table in Rmarkdown doc when knitting using `knitr::kable()` package. never have to make your own tables by hand!!! 


## Grouped Mutate

```{r}
weather_df |>
  group_by(name) |>
  mutate(mean_tmax = mean(tmax, na.rm = TRUE),
         centered_tmax = tmax - mean_tmax) |> 
  ggplot(aes(x = date, y = centered_tmax, color = name)) + 
  geom_point() 
```

```{r}
weather_df |> 
  group_by(name, month) |> 
  mutate(tmax_rank = min_rank(desc(tmax))) |> 
  filter(tmax_rank < 2)
```

lags

```{r}
weather_df |> 
  group_by(name) |> 
  mutate(yesterday_tmax = lag(tmax, 3))
```

^ `lags` are used to compare an observation to it’s previous value. This is useful, for example, to find the day-by-day change in max temperature within each station over the year:


- must use `group_by` when using `lag` so it knows where to pull the correct data from


```{r}
weather_df |> 
  group_by(name) |> 
  mutate(temp_change = tmax - lag(tmax))
```


```{r}
weather_df |>
  group_by(name) |>
  mutate(temp_change = tmax - lag(tmax)) |>
  summarize(
    sd_temp_change = sd(temp_change, na.rm = TRUE),
    temp_change_max = max(temp_change, na.rm = TRUE))
```

^ standard deviation of the temperature change across three different weather stations, and the maximum temperature in each of these locations
















