---
title: "Viz with gplot"
output: github_document
date: "2023-09-28"
---

```{r}
library(tidyverse)
library(ggridges)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```
^ use to set all figures at the same size or aspect ratio

Get the data for plotting: 

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())

weather_df
```

Make a scatterplot of precipitation across tmin and tmax:

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()
```

Pipes and things - start with this to add on more later

```{r}
weather_df |> 
  filter(name == "CentralPark_NY") |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()

ggp_nyc_weather = 
  weather_df |> 
  filter(name == "CentralPark_NY") |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```

^ `ggp_nyc_weather + ` to add on other ggplots on top of a base image or condense code

## Fancy Plot

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = 0.3) + 
  geom_smooth(se = FALSE)
```

^ `alpha` = 0.03 -> geom_points are 30% opaque and 70% transparent


Plot with Facets

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = 0.3) + 
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```

new plot using the same data!!! :)

```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = 0.3) + 
  geom_smooth() + 
  facet_grid(. ~ name)
```

assigning specific colors (go with default colors for now - ggplot has specific colors available!! look online) 

```{r}
weather_df |> 
  filter(name != "CentralPark_NY") |> 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(alpha = .7, size = .5)

ggplot(weather_df, aes(x = tmax, y = tmin)) + 
  geom_hex()
```

`geom_`hex()`, ``geom_bin2d()`, or `geom_density2d()` to show density in the data

## univariate plotting

histogram

```{r}
ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_histogram(position = "dodge", binwidth = 2)
```

density plot

```{r}
ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_density(alpha = .3, adjust = 2)
```

using boxplots

```{r}
ggplot(weather_df, aes(y = tmax, x = name)) + 
  geom_boxplot()
```

violin plots?

```{r}
ggplot(weather_df, aes(y = tmax, x = name)) + 
  geom_violin()
```
ridge plot

```{r}
ggplot(weather_df, aes(x = tmax, y = name)) + 
  geom_density_ridges()
```

`geom_line`

```{r}
weather_df |> 
  filter(name == "Mookai_HI") |> 
  ggplot(aes(x = date, y = tmax)) + 
  geom_line(alpha = 0.5) + 
  geom_point(size = 0.5)
```

^line plots are helpful for longitudinal data

## saving and embedding plots

```{r}
ggp_weather = 
    weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()

ggp_weather

ggsave("results/ggp_weather.pdf", ggp_weather)
```

change scale of plot 

```{r, fig.width = 6, fig.asp = .6}
ggp_weather
```





