---
title: "Viz with gplot II"
output: github_document
date: "2023-09-28"
---

```{r}
library(tidyverse)
library(patchwork)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```

^ what does `knitr`do again?? 

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2023-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

This results in a dataframe with `r nrow(weather_df)` observations on six variables. 


## Same plot from last time

# make a plot with labeled axes, title, legend, scale, and colors.

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Min daily temp (Degrees C)",
    y = "Max daily temp",
    color = "Location",
    caption = "Max vs min daily temp in three locations; data from rnoaa"
  ) + 
  scale_x_continuous(
    breaks = c(-15, 0, 15),
    labels = c("-15 C", "0", "15")
  ) +
  scale_y_continuous(
    trans = "sqrt", 
    position = "right")
```

^ how do you know how to pull in  `color = "Location"`???

^ can add `filter(tmax >= 20, tmax <= 30) |> ` above line 42, `ggplot` to filter data

# add `scale_colors` vs `viridis(discrete)`, and themes

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Min daily temp (Degrees C)",
    y = "Max daily temp",
    color = "Location",
    caption = "Max vs min daily temp in three locations; data from rnoaa"
  ) +
  viridis::scale_color_viridis(discrete = TRUE) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

## NOTE: use the following code chunk in the begginning of the Rmarkdown code to set all figures in the same formatting:

```{r}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## data argument ... 

```{r}
weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point() + 
  geom_smooth()

nyc_weather_df =
  weather_df |> 
  filter(name == "CentralPark_NY")

molokai_df = 
  weather_df |> 
  filter(name == "Molokai_HI")

ggplot(nyc_weather_df, aes(x = date, y = tmax, color= name)) + 
  geom_point() + 
  geom_line(data = molokai_df) 
```

^ use to describe a complex summary of the data. 

## `patchwork`

`patchwork` is designed to put scatterplots together using `ggplot`

```{r}
weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point() + 
  facet_grid(. ~ name)
```

```{r}
ggp_temp_scatter = 
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  theme(legend.position = "none")

ggp_prcp_density =
  weather_df |> 
  filter(prcp > 25) |> 
  ggplot(aes(x = prcp, fill = name)) + 
  geom_density(alpha = .5) +
  theme(legend.position = "none")

ggp_tmax_date = 
  weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  theme(legend.position = "bottom")

(ggp_temp_scatter + ggp_prcp_density) / ggp_tmax_date
```

## data manipulation

changing the factor variables to the structure that we care about using `fct_relevel`, `fct_reorder`. 

```{r}
weather_df |> 
  mutate(
    name = fct_relevel(name, c("Molokai_HI", "CentralPark_NY", "Waterhole_WA"))
  ) |> 
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()

weather_df |> 
  mutate(
    name = fct_reorder(name, tmax)
  ) |> 
  ggplot(aes(x = name, y = tmax, fill = name)) +
  geom_violin()
```


## data tidying, manipulation, AND plotting using `FAS litters` and `FAS pups` data

```{r}
litters_df = 
  read_csv("./data/FAS_litters.csv") |>
  janitor::clean_names() |>
  separate(group, into = c("dose", "day_of_tx"), sep = 3)

pups_df = 
  read_csv("./data/FAS_pups.csv") |>
  janitor::clean_names() |>
  mutate(
    sex = 
      case_match(
        sex, 
        1 ~ "male", 
        2 ~ "female"))

fas_df = 
  left_join(pups_df, litters_df, by = "litter_number")


fas_df |> 
  select(dose, day_of_tx, starts_with("pd")) |> 
  pivot_longer(
    pd_ears:pd_walk,
    names_to = "outcome", 
    values_to = "pn_day"
  ) |> 
  drop_na() |> 
  mutate(outcome = fct_reorder(outcome, pn_day)) |> 
  ggplot(aes(x = dose, y = pn_day)) + 
  geom_violin() +
  facet_grid(day_of_tx ~ outcome)
```























